Effectiveness of visual test in linear regression diagnositics

### 2022-07-27 paper_comparison/paper_comparison/paper_comparison.tex
udpate paper

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}

@@

\title{Why shouldn't you use numerical tests to diagnose the linear
regression models?}

@@

Regression analysis is a field of study with at least a hundred years of
history, and regression diagnostics is one of the essential steps in
regression analysis. The diagnostic procedure conventionally involves
evaluating the fitness of the proposed model, detecting the presence of
influential observations and outliers, checking the validity of model
assumptions and many more. In terms of diagnostic techniques, data
plots, hypothesis testing, and summary statistics are vital tools for a
systematic and detailed examination of the regression model
\citep{mansfield1987diagnostic}.

Many of those regression diagnostic methods and procedures are mature
and well-established in books first published in the twentieth century,
such as \citet{draper_applied_2014},
\citet{montgomery_introduction_2012}, \citet{belsley_regression_1980},
\citet{cook_applied_1999} and \citet{cook1982residuals}. Regardless of
the level of difficulty, one will find the importance and usefulness of
diagnostic plots being emphasized in those books repeatedly. Checking
diagnostic plots is also the recommended starting point for validating
model assumptions such as normality, homoscedasticity and linearity
\citep{anscombe_examination_1963}.

@@

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As suggested by \citet{cook1982residuals},
these kinds of diagnostic plots are commonly used to identify patterns
that are indicative of nonconstant error variance or non-linearity. Raw
residuals and studentized residuals are the two most frequently used
residuals in standard residual plots. The debt on which type of
residuals should be used is always present. While raw residuals are the
most common computer regression software package output, by applying a
scaling factor, the ability to reveal nonconstant error variance in
standard residual plots will often be enhanced by studentized residuals
in small sample size \citep{gunst2018regression}. As a two-dimensional
representation of a model in a \(p\)-dimensional space, standard
residual plots project data points onto the variable of the horizontal
axis, which is a vector in \(p\)-dimensional space. Observations with
the same projection will be treated as equivalent as they have the same
abscissa. Therefore, standard residual plots are often useful in
revealing model inadequacies in the direction of the variable of the
horizontal axis but could be inadequate for detecting patterns in other
directions, especially in those perpendicular to the variable of the
horizontal axis. Hence, in practice, multiple standard residual plots
with different horizontal axes will be examined
\citep{cook1982residuals}. Overlapping data points is a general issue in
scatter plots not limited to standard residual plots, which often makes
plots difficult to interpret because visual patterns are concealed.
Thus, for a relatively large sample size, \citet{cleveland1975graphical}
suggests the use of robust moving statistics as reference lines to give
aid to the eye in seeing patterns, which nowadays, are usually replaced
with a spline or local polynomial regression line.

@@

In addition to diagnostic plots, analysts may perform formal hypothesis
testing for detecting model defects. Depends on the alternative
hypothesis that being focused on, variety of tests can be applied. For
example, the presence of heteroskedasticity can usually be tested by
applying the White test \citep{white_heteroskedasticity-consistent_1980}
or the Breusch-Pagan test \citep{breusch_simple_1979}, which both
derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
principle and rely on the asymptotic properties of the null
distribution. For testing non-linearity, one may apply the F-test to
examine the significance of certain polynomial and non-linear forms of
the regressors, or the significance of proxy variables as in the Ramsey
Regression Equation Specification Error Test (RESET)
\citep{ramsey_tests_1969}.

As discussed in \citet{cook1982residuals}, most residual-based tests for
a particular type of departure from model assumptions are sensitive to
other types of departures. It is likely the null hypothesis is correctly
rejected but for the wrong reason, which is also known as the ``Type III
error''. Additionally, outliers will often incorrectly trigger the
rejection of the null hypothesis despite the residuals being
well-behaved \citep{cook_applied_1999}. This can be largely avoided in
diagnostic plots as experienced analysts can evaluate the acceptability
of assumptions flexibly, even in the presence of outliers.
\citet{montgomery_introduction_2012} stated that based on their
experience, statistical tests are not widely used in regression
diagnostics. The same or even larger amount of information can be
provided by diagnostic plots than the corresponding tests in most
empirical studies. Not to mention, it is almost impossible to have an
exactly correctly specified model in reality. There is a well-known
aphorism in statistics stated by George Box - ``All models are wrong,
but some are useful''. This indicates proper hypothesis tests will
always reject the null hypothesis as long as the sample size is large
enough. The outcome ``Not reject'' can be interpreted as either ``effect
size is small'' or ``sample size is small''. The outcome ``reject''
still doesn't inform us whether and how much the model defects are of
actual consequence to the inference and prediction. But still, the
effectiveness of statistical tests shall not be disrespected.
Statistical tests have a chance to provide analysts with unique
information. There are situations where no suitable diagnostic plots can
be found for a particular violation of the assumptions, or excessive
diagnostic plots need to be checked. One will have no choice but to rely
on statistical tests if there are any. A good regression diagnostic
practice should be a balanced combination of both methods.

@@

However, unlike hypothesis testing built upon rigorous statistical
procedures, reading diagnostic plots relies on graphical perception -
human's ability to interpret and decode the information embedded in the
graph \citep{cleveland_graphical_1984}, which is to some extent
subjective. Further, visual discovery suffers from its unsecured and
unconfirmed nature where the degree of the presence of the visual
features typically can not be measured quantitatively and objectively,
which may lead to over or under-interpretations of the data. One such
example is finding an over-interpretation of the separation between gene
groups in a two-dimensional projection from a linear discriminant
analysis when in fact there are no differences in the expression levels
between the gene groups and separation is not an uncommon occurrence
\citep{roy_chowdhury_using_2015}.

@@

In visual inference, a collection of test statistics
\(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where
\(\boldsymbol{\mathrm{y}}\) is the data and \(I\) is a set of all
possible visual features. \citet{buja_statistical_2009} described each
of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a
measurement of the degree of presence of a visual feature.
Alternatively, \citet{majumder_validation_2013} avoids the use of visual
features and defined the visual statistics \(T(.)\) as a mapping from a
dataset to a data plot. Both definitions of visual test statistics are
valid, but in the rest of the report the first definition will be used
as it covers some details needed by the following discussion. A visual
discovery is defined as a rejection of a null hypothesis, and the same
null hypothesis can be rejected by many different visual discoveries
\citep{buja_statistical_2009}. For regression diagnostics, the null
hypothesis would be the assumed model, while the visual discoveries
would be any findings that are inconsistent with the null hypothesis.
The same regression model can be rejected by many reasons with residual
plot, including non-linearity and heteroskedasticity as shown in Figure
@ref(fig:residual-plot-cubic-heter).

\begin{figure}
\centering
\includegraphics{paper_comparison_files/figure-latex/residual-plot-cubic-heter-1.pdf}
\caption{Residuals vs.~fitted values plot for a classical linear
regression model. The residuals are produced by fitting a two-predictor
multiple linear regression model with data generated from a cubic linear
model. From the residual plot, ``butterfly shape'' can be observed which
generally would be interpretd as evidence of heteroskedasticity.
Further, from the outline of the shape, nonlinear patterns exist. Both
visual discoveries are evidence against the null hypothesis, though
heteroskedasticity actually does not exist in the data generating
process.}
\end{figure}

@@

With the simulation of null plots being provided, another aspect of
hypothesis testing that needs to be addressed is the control of false
positive rate or Type I error. Any visual statistic
\(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value
\(c^{(i)}\) to form a hypothesis test. When a visual feature \(i\) is
discovered by the observer from a plot, the corresponding visual
statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there
is no general agreement on the measurement of the degree of presence of
a visual feature. It is only the event that
\(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed. Similarly,
if any visual discovery is found by the observer, we say, there exists
\(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\)
\citep{buja_statistical_2009}.

@@

Further, if we involve \(K\) independent observers in a visual test, and
let \(X\) be a random variable denoting the number of observers
correctly picking the actual data plot. Then, under the null hypothesis
\(X \sim \mathrm{Binom}_{K,1/m}\), and therefore, the \(p\)-value of a
lineup of size \(m\) evaluated by \(K\) observer is given as
\begin{equation} \label{eq:pvaluesingle}
P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
\end{equation}

@@

Compared to the conventional test, whose power only depends on the
parameter of interest \(\theta\), several studies
\citep[see][]{hofmann_graphical_2012, majumder_validation_2013, majumder_human_2014, roy_chowdhury_using_2015, loy_variations_2016}
have shown the power of the visual test is subject-specific. Thus, to be
able to account for individual's ability, an individual is required to
evaluate multiple lineups \citep{majumder_validation_2013}.

Suppose individuals have the same ability and a lineup has been
evaluated by multiple individuals, under the alternative hypothesis, the
estimated power for a lineup can be expressed as \(\hat{p} = x/K\), the
estimated probability of identifying the actual data plot from the
lineup. If the individual skill needs to be taken into account, and
\(L\) lineups have been evaluated by \(K\) individuals,
\citet{majumder_validation_2013} suggests that mixed effects logistic
regression model can be fit as:

\[g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},\] where \(g(.)\) is the
logit link function \(g(p) = log(p) - log(1-p)\); \(0 \leq p \leq 1\).
\(W_{li}\), \(1 \leq i \leq K\), \(1 \leq l \leq L\), is the covariate
matrix including lineup-specific elements and demographic information of
individuals, and \(\delta\) is a vector of parameters. \(Z\) is the
random effects matrix, and \(\tau\) is a vector of variables follow
\(N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})\).

Then, the estimated power for lineup \(l\) and individual \(i\) can be
calculated as
\(\hat{p}_{li} = g^{-1}(W_{li}\hat{\delta} + Z_{li}\hat{\tau}_{li})\)
\citep{majumder_validation_2013}.

@@

@@

@@

>>>>

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%

@@

\title{Effectiveness of visual test in linear regression diagnositics}

@@

Regression diagnostics is an essential step in regression analysis which
is a field of study with at least a hundred years of history. The
diagnostic procedure conventionally involves evaluating the fitness of
the proposed model, detecting the presence of influential observations
and outliers, checking the validity of model assumptions and many more.
In terms of diagnostic techniques, data plots, hypothesis testing, and
summary statistics are vital tools for a systematic and detailed
examination of the regression model \citep{mansfield1987diagnostic}.

Many of those regression diagnostic methods and procedures are mature
and well-established in books first published in the twentieth century,
such as \citet{draper1998applied}, \citet{montgomery1982introduction},
\citet{belsley_regression_1980}, \citet{cook_applied_1999} and
\citet{cook1982residuals}. Regardless of the level of difficulty, one
will find the importance and usefulness of diagnostic plots being
emphasized in those books repeatedly. Checking diagnostic plots is also
the recommended starting point for validating model assumptions such as
normality, homoscedasticity and linearity
\citep{anscombe_examination_1963}.

@@

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As suggested by \citet{cook1982residuals},
these kinds of diagnostic plots are commonly used to identify patterns
that are indicative of nonconstant error variance or non-linearity. Raw
residuals and studentized residuals are the two most frequently used
residuals in standard residual plots. The debt on which type of
residuals should be used is always present. While raw residuals are the
most common computer regression software package output, by applying a
scaling factor, the ability to reveal nonconstant error variance in
standard residual plots will often be enhanced by studentized residuals
in small sample size \citep{gunst1980regression}. As a two-dimensional
representation of a model in a \(p\)-dimensional space, standard
residual plots project data points onto the variable of the horizontal
axis, which is a vector in \(p\)-dimensional space. Observations with
the same projection will be treated as equivalent as they have the same
abscissa. Consequently, standard residual plots are often useful in
revealing model inadequacies in the direction of the variable of the
horizontal axis but could be inadequate for detecting patterns in other
directions, especially in those perpendicular to the variable of the
horizontal axis. Hence, in practice, multiple standard residual plots
with different horizontal axes will be examined
\citep{cook1982residuals}. Overlapping data points is another general
issue in scatter plots not limited to standard residual plots, which
often makes plots difficult to interpret because visual patterns are
concealed. Thus, for a relatively large sample size,
\citet{cleveland1975graphical} suggests the use of robust moving
statistics as reference lines to give aid to the eye in seeing patterns,
which nowadays, are usually replaced with a spline or local polynomial
regression line.

@@

Other than checking diagnostic plots, analysts may perform formal
hypothesis testing for detecting model defects. Depending on the
alternative hypothesis that is focused on, a variety of tests can be
applied. For example, the presence of heteroskedasticity can usually be
tested by applying the White test
\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
test \citep{breusch_simple_1979}, which are both derived from the
Lagrange multiplier test \citep{silvey1959lagrangian} principle that
relies on the asymptotic properties of the null distribution. For
testing non-linearity, one may apply the F-test to examine the
significance of specific polynomial and non-linear forms of the
regressors, or the significance of proxy variables as in the Ramsey
Regression Equation Specification Error Test (RESET)
\citep{ramsey_tests_1969}.

As discussed in \citet{cook1982residuals}, most residual-based tests for
a particular type of departure from model assumptions are sensitive to
other types of departures. It is likely the null hypothesis is correctly
rejected but for the wrong reason, which is known as the ``Type III
error''. Additionally, outliers will often incorrectly trigger the
rejection of the null hypothesis despite the residuals are well-behaved
\citep{cook_applied_1999}. This can be largely avoided in diagnostic
plots as experienced analysts can evaluate the acceptability of
assumptions flexibly, even in the presence of outliers.
\citet{montgomery1982introduction} suggested that based on their
experience, statistical tests are not widely used in regression
diagnostics. The same or even larger amount of information can be
provided by diagnostic plots than the corresponding tests in most
empirical studies. Not to mention, it is almost impossible to have an
exactly correctly specified model in reality. There is a well-known
aphorism in statistics stated by George Box - ``All models are wrong,
but some are useful''. This indicates proper hypothesis tests will
always reject the null hypothesis as long as the sample size is large
enough. The outcome ``Not reject'' can be interpreted as either ``effect
size is small'' or ``sample size is small''. and the outcome ``reject''
doesn't inform us whether and how much the model defects are of actual
consequence to the inference and prediction. But still, the
effectiveness of statistical tests shall not be disrespected.
Statistical tests have a chance to provide analysts with unique
information. In situations where no suitable diagnostic plots can be
found for a particular violation of the assumptions, or excessive
diagnostic plots are needed to be checked, one will have no choice but
to fall back on statistical tests if there exists any. A good regression
diagnostic practice should be a balanced combination of both methods.

@@

However, unlike hypothesis testing built upon rigorous statistical
procedures, reading diagnostic plots relies on graphical perception -
human's ability to interpret and decode the information embedded in the
graph \citep{cleveland_graphical_1984}, which is to some extent
subjective and indecisive. Further, visual discovery suffers from its
unsecured and unconfirmed nature where the degree of the presence of the
visual features typically can not be measured quantitatively and
objectively, which may lead to over or under-interpretations of the
data. One such example is finding an over-interpretation of the
separation between gene groups in a two-dimensional projection from a
linear discriminant analysis when in fact there are no differences in
the expression levels between the gene groups and separation is not an
uncommon occurrence \citep{roy_chowdhury_using_2015}.

@@

In visual inference, a collection of test statistics
\(T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)\) is defined, where
\(\boldsymbol{\mathrm{y}}\) is the data and \(I\) is a set of all
possible visual features. \citet{buja_statistical_2009} described each
of the test statistics \(T^{(i)}(\boldsymbol{\mathrm{y}})\) as a
measurement of the degree of presence of a visual feature.
Alternatively, \citet{majumder_validation_2013} avoids the use of visual
features and defined the visual statistics \(T(.)\) as a mapping from a
dataset to a data plot. Both definitions of visual test statistics are
valid, but in the rest of the paper the first definition will be used as
it covers some details needed by the following discussion. A visual
discovery is defined as a rejection of a null hypothesis, and the same
null hypothesis can be rejected by many different visual discoveries
\citep{buja_statistical_2009}. For regression diagnostics, the null
hypothesis would be the assumed model, while the visual discoveries
would be any findings that are inconsistent with the null hypothesis.
The same regression model can be rejected by many reasons with residual
plot, including non-linearity and heteroskedasticity as shown in Figure
\ref{fig:residual-plot-cubic-heter}.

\begin{figure}
\centering
\includegraphics{paper_comparison_files/figure-latex/residual-plot-cubic-heter-1.pdf}
\caption{Residuals vs.~fitted values plot for a classical linear
regression model. The residuals are produced by fitting a two-predictor
multiple linear regression model with data generated from a cubic linear
model. From the residual plot, ``butterfly shape'' can be observed which
generally would be interpretd as evidence of heteroskedasticity.
Further, from the outline of the shape, nonlinear patterns exist. Both
visual discoveries are evidence against the null hypothesis, though
heteroskedasticity actually does not exist in the data generating
process. \label{fig:residual-plot-cubic-heter}}
\end{figure}

@@

With the simulation mechanism of null plots being provided, another
aspect of hypothesis testing that needs to be addressed is the control
of false positive rate or Type I error. Any visual statistic
\(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value
\(c^{(i)}\) to form a hypothesis test. When a visual feature \(i\) is
discovered by the observer from a plot, the corresponding visual
statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there
is no general agreement on the measurement of the degree of presence of
a visual feature. It is only the event that
\(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed. Similarly,
if any visual discovery is found by the observer, we say, there exists
\(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\)
\citep{buja_statistical_2009}.

@@

Further, a visual test can involve \(K\) independent observers. Let
\(D_i = \{0,1\}\) be a binomial random variable denoting whether subject
\(i\) correctly detecting the actual data plot, and
\(X = \sum_{i=1}^{K}X_i\) be the number of observers correctly picking
the actual data plot. Then, by imposing a relatively strong assumption
on the visual test that all \(K\) evaluations are fully independent,
under the null hypothesis, \(X \sim \mathrm{Binom}_{K,1/m}\). Therefore,
the \(p\)-value of a lineup of size \(m\) evaluated by \(K\) observer is
given as \begin{equation} \label{eq:pvaluesingle}
P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},
\end{equation}

@@

As pointed out by \citet{vanderplas2021statistical}, though equation
(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
possible dependencies in the visual test due to repeated evaluations of
the same lineup. And it is inapplicable to visual test where subjects
are asked to select one or more ``most different'' plots from the
lineup. They summarized three common different scenarios in visual
inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
\(K\) lineups with different null plots but the same actual data plot
are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
subjects. Out of these three scenarios, Scenario 3 is the most common in
previous studies as it puts the least constraints on the experimental
design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, they decided to add \(1/s_j\) to
\(c_i\) instead of one, where \(s_j\) is the number of plots subject
\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\).

The full model was a Dirichlet-multinomial mixture distribution

\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
\end{split}\end{align}

Since the p-value calculation only needs to concern the number of times
the actual data plot being selected denoted by \(C_i\), they showed the
model can be simplified to a beta-binomial mixture distribution

\begin{align} \label{eq:beta-binomial}\begin{split}
\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
C_i&|\theta_i \sim Binomial(K, \theta_i).
\end{split}\end{align}

Thus, the visual p-value followed by the beta-binomial model is given as

\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
\end{equation}

where \(B(.)\) is the beta function defined as

\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
\end{equation}

Note that the use of equation (\ref{eq:pvalue-beta-binomial}) requires
the estimation of \(\hat{\alpha}\), which largely depends on the null
model, the type of the plot and other aesthetic features. They suggested
to estimate \(\hat{\alpha}\) visually based on the selections of null
plots of the experimental data, or to estimate \(\hat{\alpha}\)
numerically based on several additional Rorschach lineups, which is a
type of lineup containing only null plots. However, when the number of
null models are large, it could be expensive to manually estimate each
\(\alpha\) or include additional Rorschach lineups in the experiment.

Instead, in the experiments that will be described in section
\ref{experimental-design}, we adopt a simpler model implicitly used by
the \texttt{pmulti()} function of the \texttt{vinference} \texttt{R}
package. We assume the attractiveness of the plot \(i\) modelled as
\(w_i \sim Uniform(0,1)\) for \(i=1,..,m\). Let
\(\theta_i = w_i/\sum_{i=1}^{m}w_i\) be the probability of plot \(i\)
being selected by a subject. Then, given the number of selections
\(s_j\), for \(j=1,...,K\), the distribution of \(C_i\) can be
approximated by simulating the random selection process with computer.
The simulated visual test p-value is formulated as

\begin{equation} \label{eq:p-value-multi}
\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
\end{equation}

@@

The effectiveness of visual inference has already been validated by
\citet{majumder_validation_2013} under relatively simple classical
normal linear regression model settings with only one or two regressors.
Their results suggests visual test is capable of testing the
significance of a single regressor with a similar power as a t-test,
though they expressed that in general it is unnecessary to use visual
inference if there exists a conventional test and they didn't expect the
visual test to perform equally well as the conventional test. In their
third experiment, where there does not exist a proper conventional test,
visual test outperforms the conventional test for a large margin. This
is encouraging as it promotes the use of visual inference in border
field of data science where there are no existing statistical testing
procedures. In fact, lineup protocol has been integrated into some model
diagnostic tools such as \citet{loy2013diagnostic}.

With our knowledge, what haven't been examined so far is the
effectiveness of visual test relative to the equivalent conventional
test in regression diagnostics. Particularly, its ability to detect
non-linearity and heteroskedasticity compared to F-test and BP-test.

Two experiments were conducted to evaluate the effectiveness of visual
test relative to the equivalent conventional test in regression
diagnostics. The first experiment has ideal scenario for conventional
testing, where we would not expect the visual test to do better than the
conventional test. The second experiment is a scenario where the
conventional test is an approximate test, which visual test may have a
chance to match the performance of the conventional test.

As suggested by many previous studies {[}many studies refs here{]}, it
is clear some care is needed to set up the human subjects experiments.
It is best for an subject to evaluate a block of 20 lineups with varying
difficulty, with a reasonable number of ``easy'' lineups. The
explanations about each experiment includes an explantion of how the
lineups were sampled and provided to the subjects.

Participants for all the experiments were recruited through Prolific's
online web service {[}ref here{]}. A summary of the data obtained for
all three eperiments are shown in {[}table ref here{]}. Participants
were asked to select the plots which are mostly different from others,
provide a reason for their selections, and express how different they
think the selected plots are from others. Gender, age, education, and
previous experience in visual experiment are also collected. For each of
the experiments, \(12\) of the lineups were used as a attention check
which everyone will randomly get two of them and should get at least one
correct, so that a measure of the quality of the subjects effort could
be made. Note that no participant was shown the same lineup twice.

\hypertarget{non-linear-covariates}{%
\subsection{Non-linear covariates}\label{non-linear-covariates}}

The experiment is designed to study the ability of human subjects to
detect the effect of a higher order term \(Z\) in a two variable
statistical model formulated as:

\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{Y} = 1 + \boldsymbol{X} + \boldsymbol{Z} + \boldsymbol{\varepsilon},\\
\boldsymbol{X} = g(\boldsymbol{X_{raw}}, 1), \\
\boldsymbol{Z} = g(\boldsymbol{Z_{raw}}, 1), \\
\boldsymbol{Z_{raw}} = He_j(g(\boldsymbol{X}, 2)),
\end{align}

where \(\boldsymbol{Y}\), \(\boldsymbol{X}\),
\(\boldsymbol{\varepsilon}\), \(\boldsymbol{X_{raw}}\),
\(\boldsymbol{Z_{raw}}\) are \(n\times1\) matrices, \(He_{j}(.)\) is the
\(j\)th-order probabilist's Hermite polynomials,
\(j \in \{2, 3, 6, 18\}\),
\(\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\),
\(\sigma \in \{0.5, 1, 2, 4\}\), and \(g(\boldsymbol{X}, k)\) is a
scaling function to enforce the support of the variable to be
\(\{-k, k\}\) defined as

\begin{equation} \label{eq:scaling-function}
g(\boldsymbol{X}, k) = (\boldsymbol{X} - min(\boldsymbol{X}))/max(\boldsymbol{X} - min(\boldsymbol{X})) \times 2k - k, \quad \text{for} \quad k > 0. 
\end{equation}

The null regression model used to fit the data genearted by the above
model is formulated as:

\begin{equation} \label{eq:scaling-function}
\boldsymbol{Y} = \beta_0 + \beta_1 \boldsymbol{X} + \boldsymbol{u},
\end{equation}

where
\(\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)\).

Clearly, omitted-variable bias will present since the null model leaves
out the high order term. Data were simulated using a range of values of
\(j\) as shown in {[}table ref here{]}, three different smaple sizes (n
= 50, 100, 300) and four different standard deviations of the error
(\(\sigma\) = 0.5, 1, 2, 4). The range of \(j\) was chosen so that
different shapes of non-linearity are included in the residual plot.

The range of \(sigma\) was chosen so that estimates of the power would
produce reasonably continuous power curves, comparable to that
calculated for the conventional test. The values for \(X_{raw}\) were
randomly generated independently from one of the following four
distributions: (1) \(U(a = -1, b = 1)\), (2)
\(N(\mu = 0, \sigma = 0.3)\), (3)
\(lognormal(\mu = 0, \sigma = 0.6)/3 - 1\) and (4)
\(u\{a = 1, b = 5\}\).

@@

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-3-1.pdf}

@@

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-4-1.pdf}
